Escaping RGBland: Selecting Colors for Statistical Graphics

### 2008-10-23 pkg/paper/Palettes/CSDA.Rnw
KH feedback
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@47 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

A color model typically supported by software packages involves the specification
of colors as Red-Green-Blue (RGB) triplets. However, this specification
corresponds to color generation on a computer screen \citep[see][]{color:Poynton:2000}
rather than corresponding to human color perception. For humans, it is
virtually impossible to control the perceptual properties of a color in this
color space because there is no single dimension that corresponds to, e.g.,
the hue or the brightness of the color. As a consequence, various perceptually-based
color spaces have been suggested, where each dimension of the color space can
be matched with a perceptual property. One approach widely implemented in
software packages involves Hue-Saturation-Value (HSV) triplets \citep{color:Smith:1978},
a simple transformation of RGB triplets \citep[see][]{color:Wiki+HSV:2007}.
Unfortunately, the dimensions in HSV space map poorly to perceptual
properties and the use of HSV colors encourages the use of 
highly saturated colors. A perceptually based color model that mitigates
these problems involves Hue-Chroma-Luminance (HCL) triplets \citep[see][]{color:Ihaka:2003},
which is a transformation of CIELUV color space \citep[see][]{color:Wiki+CIELUV:2007}.
This is the color space we advocate in this article.

@@

The remainder of the paper is organized as follows: Section~\ref{sec:motivation} provides
some motivating examples, showing how typical HSV-based graphics can be enhanced
by using HCL-based palettes. Section~\ref{sec:spaces} gives a brief introduction
to the underlying HSV and HCL color spaces before Section~\ref{sec:palettes} 
suggests strategies for deriving HCL-based qualitative, sequential, and diverging palettes.
Some more more illustrations are presented in Section~\ref{sec:illustrations}.
Section~\ref{sec:software} offers some general remarks on the implementation in statistical
software packages as well as some details on the implementation in the \proglang{R}
system for statistical computing \citep{color:R:2008}.
Section~\ref{sec:discussion} concludes the paper with a discussion.

@@

%% \section*{Acknowledgments}
%% 
%% We are thankful to David Meyer, Ross Ihaka, Ken Knoblauch, Brian~D.~Ripley,
%% Thomas Kneib, Michael H\"ohle, and Thomas Lumley for feedback, suggestions and discussions.

>>>>

A color model typically supported by software packages involves the specification
of colors as Red-Green-Blue (RGB) triplets. However, this specification
corresponds to color generation on a computer screen \citep[see][]{color:Poynton:2000}
rather than corresponding to human color perception. For humans, it is
virtually impossible to control the perceptual properties of a color in this
color space because there is no single dimension that corresponds to, e.g.,
the hue or the brightness of the color. As a consequence, various perceptually-based
color spaces have been suggested, where each dimension of the color space can
be matched with a perceptual property. One approach widely implemented in
software packages involves Hue-Saturation-Value (HSV) triplets \citep{color:Smith:1978},
a simple transformation of RGB triplets \citep[see][]{color:Wiki+HSV:2007}.
Unfortunately, the dimensions in HSV space map poorly to perceptual
properties and the use of HSV colors encourages the use of 
highly saturated colors. A perceptually based color model that mitigates
these problems involves Hue-Chroma-Luminance (HCL) triplets \citep[see][]{color:Ihaka:2003},
resulting from a transformation of CIELUV color space \citep[see][]{color:Wiki+CIELUV:2007}.
This is the color space we advocate in this article.

@@

The remainder of the paper is organized as follows: Section~\ref{sec:motivation} provides
some motivating examples, showing how typical HSV-based graphics can be enhanced
by using HCL-based palettes. Section~\ref{sec:spaces} gives a brief introduction
to the underlying HSV and HCL color spaces before Section~\ref{sec:palettes} 
suggests strategies for deriving HCL-based qualitative, sequential, and diverging palettes.
Some more illustrations are presented in Section~\ref{sec:illustrations}.
Section~\ref{sec:software} offers some general remarks on the implementation in statistical
software packages as well as some details on the implementation in the \proglang{R}
system for statistical computing \citep{color:R:2008}.
Section~\ref{sec:discussion} concludes the paper with a discussion.

@@

\section*{Acknowledgments}

We are thankful to Michael H\"ohle, Ross Ihaka, Thomas Kneib, Ken Knoblauch,
Thomas Lumley, David Meyer, and Brian~D.~Ripley for feedback, suggestions, and discussions.
